### 2024-11-25 paper/paper.qmd
Paper draft over.

---
title: "Pathways to Early Vocabulary Acquisition"
subtitle: "Predicting English Language Development Trajectories in Children Aged 16–30 Months"
author: Yongqi Liu
thanks: "Code and data are available at: https://github.com/Cassieliu77/Vocabulary_Learning_Pattern.git"
date: today
date-format: long
abstract: "This study examines vocabulary acquisition in children aged 16–30 months using logistic regression models and data from the MacArthur-Bates Communicative Development Inventories. Age is the strongest predictor, with older children showing higher vocabulary proficiency, while foundational word categories like Function Words are acquired more consistently than complex ones such as Sensory Words. Children in norming groups outperform non-norming peers, highlighting the impact of structured linguistic environments. The findings support targeted interventions to address variability in complex vocabulary acquisition and promote balanced language development."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

# List of required packages
# packages <- c("tidyverse", "arrow", "modelsummary", "knitr", "dplyr", 
#            "ggplot2", "scales", "wordbankr", "tidyr", "arm")

library(tidyverse)
library(arrow)
library(modelsummary)
library(knitr)
library(dplyr)
library(ggplot2)
library(scales)
library(wordbankr)
library(tidyr)
library(arm)

analysis_data <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
logistic_model <- readRDS(here::here("models/logistic_model_last.rds"))
```

# Introduction {#sec-intro}

Understanding how children acquire vocabulary is a cornerstone of developmental linguistics and early childhood education. Vocabulary growth is not only a key indicator of cognitive development but also serves as a foundation for future linguistic and academic success. The interplay of factors such as age, linguistic environments, and word categories influences children’s vocabulary acquisition in complex and dynamic ways. By analyzing patterns and predictors of vocabulary growth, researchers can illuminate the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities.

This study utilizes a dataset from the @wordbankr database, which aggregates data from the MacArthur-Bates Communicative Development Inventories (CDI), to investigate vocabulary development in children aged 16–30 months. By employing a logistic regression model, the research estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories. This model serves as a tool to explore how these factors collectively influence children’s vocabulary acquisition trajectories.

The estimand of interest in this study is the conditional probability that a child possesses advanced vocabulary proficiency, defined as a high combined comprehension and production score, given specific predictors. These predictors include the child’s age in months, their norming status (whether they belong to a standardized sample used for benchmarking), and the lexical category of the words (e.g., Function Words, Verbs, Adjectives). By employing a logistic regression model, the research estimates how these variables individually and interactively influence the likelihood of achieving advanced vocabulary levels. The model demonstrates developmental trends, such as how median vocabulary proficiency increases with age, and reveals how specific word categories contribute uniquely to language growth. For example, foundational categories like Function Words and Objects may show consistently higher probabilities of acquisition, while more abstract categories, such as Verbs and Sensory Words, exhibit greater variability due to their complexity or context-specific nature.

Key findings reveal distinct trajectories of vocabulary acquisition across different word categories, with age emerging as the strongest predictor of vocabulary growth. The analysis demonstrates that as children age, their likelihood of achieving high vocabulary scores consistently increases, reflecting the natural progression of language development. Categories such as “Living Things” and “Function Words” display stable and high predicted probabilities, likely due to their frequent occurrence in everyday communication and their role in forming early linguistic structures. Conversely, categories such as “Sensory Words” and “Adjectives” show more variability, highlighting the challenges posed by their later acquisition and the contextual understanding they require.

Additionally, children in the norming group, who are assessed in structured linguistic environments, consistently achieve higher vocabulary scores compared to non-norming children. This difference underscores the influence of systematic linguistic exposure in shaping vocabulary development. The results not only provide a nuanced understanding of vocabulary acquisition but also emphasize the importance of targeted language interventions. They suggest that focusing on underdeveloped categories, such as Sensory Words and Adjectives, could help bridge gaps in vocabulary learning. Furthermore, these findings set the stage for future research into the broader contextual factors, such as socio-economic status and bilingualism, that may further explain variability in language acquisition trajectories.

The remainder of this paper is structured as follows: @sec-intro provides an introduction and decribe the estimand for the paper. @sec-data1 describes the dataset, highlighting the variables used and their distributions. @sec-model1 provides an overview of the logistic regression model and its assumptions. @sec-result presents the visualized results, including variability in vocabulary acquisition, the role of predictors, and predicted probabilities. Finally, @sec-discussion discusses the implications of the findings and outlines limitations and directions for future research. The data used in this study all comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR.

# Data {#sec-data1}

## Overview

The original dataset was obtained from @wordbankr. After undergoing a thorough cleaning process—including grouping related items and removing missing values—the analysis focuses on the key variables: category, age, comprehension, production, is_norming, and broad_category. These variables form the foundation of the analysis dataset. An overview of the cleaned dataset is presented in @tbl-data.

```{r}
#| label: tbl-data
#| tbl-cap: Cleaned Word Bank Dataset
#| echo: false

summary_table <- analysis_data %>%
  dplyr::select(language, age, is_norming, broad_category, production, high_vocabulary) %>%
  rename(
    Language = language,
    Age = age,
    Is_Norming = is_norming,
    Broad_Category = broad_category,
    Production = production,
    High_Vocabulary = high_vocabulary)

kable(head(summary_table, 10))
```

## Measurement

### Data Collection and Quality Control

The objective of measurement in this study is to translate raw parental reports into reliable indicators of vocabulary acquisition patterns in children. The data is derived from the MacArthur-Bates Communicative Development Inventories (CDI), a widely used tool that collects information on children’s vocabulary comprehension and production through structured parental surveys. These surveys allow parents to report on their child’s understanding and use of specific words, grouped into lexical categories such as nouns, verbs, and adjectives. The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
- Structured Response Formats: The CDI employs predefined response options for comprehension and production, reducing ambiguity and enhancing consistency in parental reporting.
- Words Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
- Norming Group Comparison: The inclusion of norming groups as benchmarks helps to ensure the validity of reported vocabulary scores and allows for cross-child comparison. These groups provide a standardized reference for analyzing individual differences in language acquisition.
- Variable Standardization: Continuous variables, such as age, are standardized (e.g., scaled) to reduce variability and improve the interpretability of statistical models. This ensures that coefficients reflect meaningful changes in relation to standardized measures.
- Bias Mitigation: By structuring responses and including norming benchmarks, the CDI minimizes some of the biases inherent in self-reported data, such as over- or underestimation by parents.
- Missing Data Handling: Observations with incomplete or invalid responses were excluded from the analysis to maintain the integrity and reliability of the dataset.

### Reporting Bias

However, there are several considerations regarding the data collection process:

-	Parental Reporting Bias: The reliance on parental reports introduces the potential for bias, including overestimation or underestimation of a child’s abilities. This is inherent to self-reported data and can affect the accuracy of the results.
-	Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents. This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
-	Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison. While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
-	Temporal Limitations: The CDI data represents snapshots of vocabulary development at specific ages, which may not account for rapid changes or variations over time in a child’s language acquisition process.

Despite its standardized structure, the CDI is subject to biases inherent in parental reporting, including over or underestimation bias. Parents may unintentionally overestimate or underestimate their child’s skills due to subjective perceptions or limited observations. Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.

## Outcome Variable

### High Vocabulary Score

The outcome variable in this study, High Vocabulary Score, is a binary indicator designed to identify individuals with advanced vocabulary proficiency. This variable is derived from two key measures:

1. Comprehension: This variable represents the ability to understand words and phrases, reflecting the receptive language skills of individuals. Comprehension scores are numerical and vary across the dataset.

2. Production: This variable captures the ability to produce words, reflecting expressive language skills. Like comprehension, production scores are numerical and provide the standard into verbal articulation capabilities.

3. The High Vocabulary Score is calculated using the average of comprehension and production scores for each individual. This average is represented as: $\text{prod\_comp\_mean} = \frac{\text{Comprehension} + \text{Production}}{2}$

Based on @wendy, I looked into the normal vocabulary size change among the children aged 16 to 30 months. To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
- Individuals with `prod_comp_mean > 300` are classified as having a high vocabulary score (outcome = 1).
- Those with `prod_comp_mean <= 300` are classified as not having a high vocabulary score (outcome = 0).

This approach ensures that both receptive (comprehension) and expressive (production) skills are considered in defining advanced vocabulary. The threshold of 300 was chosen based on exploratory analysis of the dataset, reflecting a meaningful distinction between individuals with high and low vocabulary abilities. The High Vocabulary Score serves as the dependent variable in the following data analysis part. Its binary nature makes it suitable for modeling with a binomial family distribution, allowing for the estimation of factors that influence advanced vocabulary acquisition.

```{r}
#| label: fig-outcome
#| fig-cap: Distribution of the outcome variable, showing the counts of children classified as having “High Vocabulary” and “Not High Vocabulary” based on their comprehension and production scores. The bar plot illustrates the balance between the two categories in the dataset, which is important for modeling purposes.
#| echo: false
#| warning: false
#| message: false

outcome_counts <- analysis_data %>%
  count(high_vocabulary) %>%
  mutate(
    VocabularyCategory = ifelse(high_vocabulary == 1, "High Vocabulary", "Not High Vocabulary"))

ggplot(outcome_counts, aes(x = VocabularyCategory, y = n, fill = VocabularyCategory)) +
  geom_bar(stat = "identity", width = 0.7, show.legend = FALSE) +
  labs(
    title = "Distribution of High Vocabulary Levels",
    x = "Vocabulary Level",
    y = "Count") +
  scale_fill_manual(values = c("High Vocabulary" = "blue", "Not High Vocabulary" = "red")) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    axis.text = element_text(size = 12),)

```

\newpage

## Predictor variables

### Age

@fig-age displays the distribution of children’s ages (in months) within the dataset, highlighting key patterns in the sample’s demographic structure. A notable concentration of data is observed among children aged between 24 and 30 months, reflecting an emphasis on capturing vocabulary development during critical periods of language acquisition. These age ranges are known to mark significant milestones in linguistic growth, which could explain their higher representation. Conversely, younger age groups (below 20 months) are underrepresented, likely due to the challenges of assessing vocabulary at earlier stages of development, where verbal communication is less pronounced and parental reporting is more variable.

The dataset also shows distinct peaks at specific ages, such as 25 and 30 months. These sharp spikes may reflect intentional focus points for testing or developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI). This uneven age distribution underscores the importance of age as a critical factor in analyzing vocabulary acquisition. While the high concentration of data at older ages enhances the analysis into advanced vocabulary development, it also necessitates caution in generalizing findings to underrepresented age groups. This observation emphasizes the need to standardize age in statistical models to account for variability across different age groups.

```{r}
#| echo: false
#| label: fig-age
#| fig-cap: It shows the distribution of children’s ages (in months) within the dataset. The majority of observations fall between 24 and 30 months, with noticeable peaks at 25 and 30 months, reflecting a focus on key developmental periods. Younger age groups are underrepresented, highlighting the need to account for variability in age when analyzing vocabulary acquisition patterns.
#| warning: false
#| message: false

ggplot(analysis_data, aes(x = age)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.8) +
  labs(title = "Age Distribution", x = "Age (months)", y = "Count") +
  theme_minimal()+
  scale_y_continuous(labels = scales::comma)

```

### Lexical Category
The words in the dataset were grouped into broad lexical categories to facilitate the analysis of vocabulary acquisition patterns. These categories include Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs. The classification was based on the semantic and functional roles of words, with nouns subdivided into more specific groups such as Living Things, Objects, and Places to capture distinct trends in vocabulary acquisition. For instance, Function Words include pronouns and question words, reflecting grammatical development, while Verbs and Adjectives capture action and descriptive words, essential for sentence construction and expression.

The bar graph illustrates the distribution of items across these categories, highlighting significant variation in word frequency. Objects constitute the largest category, suggesting a focus on tangible and concrete items, which are likely easier for children to recognize and recall. This is followed by Verbs and Living Things, categories that are fundamental to communication but slightly less prevalent. In contrast, Sensory Words and Activities are sparsely represented, possibly reflecting their specialized and context-dependent nature. The distribution underscores the importance of concrete and functional words in early vocabulary development while highlighting potential gaps in underrepresented categories. This variation provides a foundation for exploring how lexical diversity influences vocabulary acquisition patterns.

```{r}
#| echo: false
#| label: fig-category
#| fig-cap: The figure shows objects dominate the vocabulary, reflecting an emphasis on concrete and tangible terms, while categories like Sensory Words and Activities are less frequently represented, indicating the relative complexity or specificity of these word types in early language acquisition
#| warning: false
#| message: false

ggplot(analysis_data, aes(x = broad_category)) +
  geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
  labs(title = "Reclassified Category Distribution", x = "Lexical Category", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(labels = scales::comma)
```

\newpage

### Norming Status

The is_norming variable categorizes children into two groups: the norming group and the non-norming group. The norming group serves as a standardized sample, providing a benchmark for assessing vocabulary development and enabling reliable comparisons across the dataset. This group is essential for ensuring the validity of the analysis by offering a consistent reference point for evaluating individual and group-level differences in vocabulary acquisition.

The dataset is predominantly composed of non-norming children, with only a small proportion belonging to the norming group. This imbalance reflects the inherent challenges of including a sufficiently broad and representative norming sample in large-scale assessments. While the non-norming group captures diverse linguistic and demographic contexts, the norming group remains vital for calibrating and interpreting vocabulary development patterns, particularly in early language acquisition studies. @tbl-norming illustrates this distribution, highlighting the predominance of non-norming children in the dataset.

The is_norming variable is included in the logistic regression model to account for potential systematic differences between the norming and non-norming groups. As the norming group represents a standardized sample used as a benchmark for vocabulary development, its inclusion ensures that variations in the likelihood of achieving a high vocabulary score are not confounded by differences in group composition or assessment protocols. It allows the model to estimate whether membership in the norming group significantly influences vocabulary acquisition outcomes. This is particularly important given the inherent imbalance in the dataset, where the non-norming group constitutes the majority. By controlling for is_norming, the model can differentiate between developmental patterns attributable to broader population diversity versus those arising from the structured selection of the norming sample.

```{r}
#| echo: false
#| label: tbl-norming
#| tbl-cap: The dataset is primarily composed of non-norming children, with a smaller subset belonging to the norming group, serving as a standardized benchmark for assessing vocabulary development
#| warning: false
#| message: false

summary <- summary(analysis_data$is_norming)
# Summarize the 'is_norming' variable
summary_isnorming <- table(analysis_data$is_norming)

# Convert the table into a data frame
summary_isnorming_df <- as.data.frame(summary_isnorming)

# Render the summary table using kable
knitr::kable(
  summary_isnorming_df,
  col.names = c("Norming Status", "Count"),  
  caption = "Summary of Norming Status in the Dataset")
```

# Model {#sec-model1}

## Model Selection

To investigate the relationship between children’s vocabulary acquisition and their demographic and linguistic characteristics, we constructed a logistic regression model. By examining key demographic and linguistic predictors, we aim to identify how characteristics like age, norming status, and word categories influence vocabulary development. The dependent variable, high_vocabulary, is a binary outcome indicating whether a child’s average production and comprehension score (denoted as prod_comp_mean) exceeds 300. This threshold was chosen to distinguish children with relatively advanced vocabulary levels. More background details and diagnostics are included in [Appendix-@sec-model2].

## Logistic Regression Model Overview

- High Vocabulary: A binary indicator where 1 represents a high vocabulary score (combined comprehension and production > 300), and 0 otherwise.

- Scaled Age (age_scaled): The child’s age, standardized to reflect changes per standard deviation. Standardization aids in interpretability and ensures numerical stability.

- Norming Status (is_norming): A binary indicator denoting whether a child is part of the norming dataset (TRUE) or not (FALSE). This variable accounts for potential differences in data collection or assessment protocols.

- Broad Category (broad_category): A categorical variable grouping words into lexical categories, such as adjectives, verbs, and nouns. The reference category for comparison is Function Words.

The model is specified as:

\begin{align}
\log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \cdot \text{age\_scaled}_i + \beta_2 \cdot \text{is\_normingTRUE}_i \\
&\quad + \beta_3 \cdot \text{broad\_categoryAdjectives}_i \\
&\quad + \beta_4 \cdot \text{broad\_categoryFunction\_Words}_i \\
&\quad + \beta_5 \cdot \text{broad\_categoryLiving\_Things}_i \\
&\quad + \beta_6 \cdot \text{broad\_categoryObjects}_i \\
&\quad + \beta_7 \cdot \text{broad\_categoryPlaces}_i \\
&\quad + \beta_8 \cdot \text{broad\_categorySensory\_Words}_i \\
&\quad + \beta_9 \cdot \text{broad\_categoryVerbs}_i
\end{align}

Where:
- $p_i$ represents the probability that child  i  has a high vocabulary score
- $\beta_0$ is the intercept, capturing the baseline log-odds when all predictors are at their reference or mean levels
- $\beta_1$: Effect of age (standardized)
- $\beta_2$: The effect of whether the individual belongs to the norming group
- $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).

## Model Assumptions

- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary proficiency) and the independent variables. For instance, the standardized age variable (age\_scaled) ensures that for every one standard deviation increase in age, the log-odds of achieving high vocabulary change by a constant amount. Standardization centers the variable around zero and scales it to have a standard deviation of one, improving both interpretability and adherence to the linearity assumption.
- Independence of Observations: Each observation corresponds to a unique child, and the data assumes no repeated measures or nested structures (e.g., grouping by classrooms or schools). This independence ensures the validity of the logistic regression framework. If dependencies, such as repeated measures or clustering, were present, a mixed-effects logistic regression or other hierarchical modeling techniques would be required.
- Categorical Variable Encoding: The categorical variable broad_category (e.g., “Adjectives,” “Verbs,” “Living Things”) is encoded using sum contrasts. This ensures that each coefficient reflects the deviation of a given category from the overall mean effect across all categories. For example, the coefficient for “Verbs” represents the difference in log-odds for this category compared to the average log-odds across all categories. Sum contrasts further allow the model’s intercept to represent the overall mean effect when all predictors are at their reference or mean levels, facilitating clear interpretation.
- Binary Nature of the Outcome: The dependent variable, high_vocabulary, is binary (1 = high vocabulary, 0 = not high vocabulary). Logistic regression is appropriate for modeling binary outcomes, as it assumes the binomial distribution of the data, which aligns with the structure of the outcome variable.
- No Perfect Multicollinearity: The predictors are assumed to be non-perfectly correlated. High multicollinearity would lead to unreliable coefficient estimates and obscure the individual effects of predictors. Standardizing continuous predictors (e.g., age) and using appropriate encoding for categorical variables help minimize this risk and ensure stable model estimation.

## Interpretation of Coefficients

The logistic regression coefficient ($\beta$) represent the change in the log-odds of achieving high vocabulary proficiency for a one-unit change in the respective predictor variable, holding all other variables constant.

- Intercept ($\beta_0$):  Represents the log-odds of high vocabulary proficiency when all predictors are at their reference levels. If $\beta_0$ > 0, the baseline odds of high vocabulary are greater than 50%. Refering to the modelsummary in the @sec-model2, we knwo that the baseline odds of high vocabulary proficiency are less than 50%.

- Scaled Age ($\beta_1$): For each one standard deviation increase in age, the log-odds of high vocabulary increase by $\beta_1$. If $\beta_1 = 0.5$, then $\exp(0.5) \approx 1.65$, meaning the odds increase by 65% for every one standard deviation increase in age.

- Norming Status ($\beta_2$): Indicates the effect of belonging to the norming group, a positive $\beta_2$ suggests higher odds of high vocabulary compared to non-norming children. If $\beta_2 = 0.1$, then $\exp(0.1) \approx 1.11$, meaning norming group children have 22% higher odds of high vocabulary.

- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$): The coefficients for the broad word categories (e.g., Function Words, Living Things) represent their effect on the log-odds of high vocabulary compared to the reference category (Adjectives). A positive $\beta_k$ indicates higher odds compared to Adjectives. For example, if $\beta_4$ = 0.3 (Function Words), then $\exp(0.3) \approx 1.35$, meaning Function Words increase the odds by 35% compared to Adjectives. A negative $\beta_k$ implies lower odds. For instance, if $\beta_7 = -0.008$ (Places), then $\exp(-0.008) \approx 0.992$ shows a minor decrease in odds for words in the Places category.

## Model Justification

Logistic regression is well-suited for this study, given its ability to model binary outcomes like high vocabulary proficiency (1 = high vocabulary, 0 = not high vocabulary) while maintaining interpretability. By constraining predicted probabilities between 0 and 1, logistic regression ensures meaningful predictions. Its coefficients offer clear investigation into the magnitude and direction of effects, enabling an understanding of how predictors like age and norming status influence the likelihood of high vocabulary acquisition. For instance, odds ratios derived from the model allow straightforward interpretation of the impact of each variable, such as the increased likelihood of high vocabulary with a one-standard-deviation increase in age.

Although more advanced machine learning models like random forests or neural networks could provide slight improvements in predictive accuracy, these approaches lack the transparency needed to understand the underlying relationships between predictors and outcomes. Given this study’s focus on uncovering developmental patterns rather than solely optimizing prediction accuracy, logistic regression offers the necessary balance between interpretability and performance. Moreover, the dataset size and structure favor logistic regression, which is less prone to overfitting than more complex models that often require larger datasets to generalize effectively.

To ensure model robustness, the dataset was split into training and testing subsets to validate the model and minimize overfitting. Additionally, predictors like age were standardized to ensure comparability and prevent dominance by variables with larger numerical ranges. Categorical variables, such as word categories, were encoded using sum contrasts, allowing meaningful comparisons and ensuring that coefficients reflect deviations from the overall mean effect. These steps, combined with logistic regression’s simplicity and explanatory power, make it an optimal choice for investigating vocabulary acquisition patterns.

# Results {#sec-result}

## Variability in Production Vocabulary

@fig-percent visualizes the relationship between age and production vocabulary scores, focusing on different percentiles of the distribution. The scatterplot shows individual production scores as gray dots, while overlaid lines represent percentiles (10th, 25th, 50th, 75th, and 90th), capturing central tendencies and variability across ages. The 50th percentile (median) line provides a benchmark for the typical vocabulary production score at each age, whereas the 10th and 90th percentiles outline the lower and upper ranges of vocabulary development. The gradual upward trend of the median line reflects consistent growth in production vocabulary as children age, with a widening gap between the percentiles at later ages. This widening suggests increasing variability in vocabulary acquisition, with some children advancing much faster than others in production abilities.

The data indicates that children in the 90th percentile acquire vocabulary at a significantly faster rate than their peers, as evidenced by the steeper slope of the topmost line. Conversely, the 10th and 25th percentiles show more gradual, stable growth, suggesting slower development for children in these groups. The broader range of scores at older ages emphasizes the heterogeneity of developmental trajectories, with some children reaching vocabulary sizes substantially larger than the median while others remain below average. These findings underscore the diversity in early language acquisition and highlight the importance of considering individual differences when evaluating children’s vocabulary development.

```{r}
#| echo: false
#| eval: true
#| label: fig-percent
#| fig-cap: Production Vocabulary by Age with Percentile Lines. The graph illustrates the production scores of children across different ages (in months). Individual data points (gray dots) represent raw production scores for each participant. Colored lines correspond to standardized percentiles—10th (blue), 25th (light blue), 50th (green), 75th (yellow), and 90th (red)—showing trends in vocabulary production distribution over time.
#| warning: false
#| message: false

# Simulate some data for demonstration purposes
set.seed(666)
analysis_data <- data.frame(
  age = runif(1000, 15, 30),
  production = rgamma(1000, shape = 2, scale = 50))

# Define the percentiles and process the data
percentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)

percentile_data <- analysis_data %>%
  mutate(age_rounded = round(age)) %>%
  group_by(age_rounded) %>%
  summarize(
    `10th` = quantile(production, probs = 0.1, na.rm = TRUE),
    `25th` = quantile(production, probs = 0.25, na.rm = TRUE),
    `50th` = quantile(production, probs = 0.5, na.rm = TRUE),
    `75th` = quantile(production, probs = 0.75, na.rm = TRUE),
    `90th` = quantile(production, probs = 0.9, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = -age_rounded,
    names_to = "percentile",
    values_to = "value")

# Visualize the data
ggplot() +
  geom_point(data = analysis_data, aes(x = age, y = production), alpha = 0.2, color = "grey") +
  geom_line(data = percentile_data, aes(x = age_rounded, y = value, color = percentile, group = percentile), size = 1) +
  scale_color_manual(
    values = c("blue", "lightblue", "green", "gold", "red"),
    labels = c("10th", "25th", "50th", "75th", "90th")
  ) +
  labs(
    title = "Production Vocabulary by Age with Percentile Lines",
    x = "Age (Months)",
    y = "Production (Number of Words)",
    color = "Percentile"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    legend.position = "top"
  )

```

## Median Vocabulary Size Change by Age

@fig-median illustrates the median comprehension vocabulary scores across different ages, focusing on the central tendency of children’s comprehension development between 15 and 30 months. The gray dots represent individual data points, capturing the variability in comprehension scores, while the blue line highlights the median score for each age group. The graph shows a clear upward trajectory, with median comprehension steadily increasing with age, particularly after 18 months. This suggests a critical developmental period between 18 and 30 months during which children experience significant growth in comprehension vocabulary. The density and spread of gray points around the median line indicate individual variability, emphasizing that while the general trend is one of growth, some children exhibit slower or faster development compared to their peers. The visualization underscores the importance of age as a determinant of vocabulary comprehension while highlighting the diverse range of learning patterns among children.

Steady Growth in Comprehension Vocabulary: Between 15 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition. A noticeable increase in vocabulary size is observed after 18 months, suggesting that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop. The most significant growth occurs between 24 and 30 months, where the median comprehension score consistently rises. This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.The distribution of points (grey scatter) highlights considerable variability in comprehension scores at each age. While the median line captures the central trend, some children show significantly higher or lower comprehension compared to their peers, reflecting individual differences in language learning rates.Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.

The use of the median instead of the mean ensures that the central trend is not skewed by outliers (e.g., extremely high or low comprehension scores). This choice provides a robust summary of comprehension at each age, especially in datasets with large variability or non-normal distributions.These findings highlight the critical window between 21 and 26 months for comprehension vocabulary growth. Interventions or language exposure strategies during this period may be particularly effective in enhancing language development. The observed variability suggests that individual-level factors (e.g., family environment, exposure to language) play a significant role in shaping comprehension scores, warranting further investigation into these influences.

```{r}
#| echo: false
#| eval: true
#| label: fig-median
#| fig-cap: This figure illustrates the relationship between age (in months) and the median comprehension vocabulary level. Each orange point represents a resampled data point, while the blue line depicts the median vocabulary level at each age. The trend highlights a steady increase in vocabulary comprehension as children age, with noticeable variability across individual data points.
#| warning: false
#| message: false

analysis_data_1 <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))

# Resample the dataset to include 100 observations
set.seed(666)  # For reproducibility
sampled_data <-  analysis_data_1 %>%
  sample_n(800)

# Calculate the median comprehension for each age in the sampled dataset
median_data <- sampled_data %>%
  group_by(age) %>%
  summarize(
    median_prod_comp_mean = median(prod_comp_mean, na.rm = TRUE), 
    .groups = "drop")

# Create a scatter plot with median comprehension by age
ggplot() +
  geom_jitter(data = sampled_data, aes(x = age, y = prod_comp_mean), 
              alpha = 0.7, size = 2, color = "orange") +  # Scatter points for sampled data
  geom_line(data = median_data, aes(x = age, y = median_prod_comp_mean), 
            color = "blue", size = 1.2, linetype = "solid") +  # Line for median comprehension
  labs(
    title = "Median Comprehension Vocabulary by Age (Resampled)", 
    x = "Age (Months)",
    y = "Median of the Vocabulary Level") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(15, 30, 3))
```

## Prediciton for the Probability of High Vocabulary Level

```{r}
#| echo: false
#| eval: true
#| label: fig-goodgood
#| fig-cap: How the probability of high vocabulary varies with age by aggregating the predictions for each age group
#| warning: false
#| message: false

# Add predicted probabilities to the test dataset
analysis_data_test<- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
analysis_data_test <- analysis_data_test %>% mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))

# Aggregate mean predicted probabilities by age and broad_category
age_broad_category_predictions <- analysis_data_test %>%
  group_by(age, broad_category, is_norming) %>%
  summarize(mean_predicted_prob = mean(predicted_prob, na.rm = TRUE), .groups = "drop")

broad_category_summary <- age_broad_category_predictions %>%
  group_by(broad_category) %>%
  summarize(overall_mean_prob = mean(mean_predicted_prob, na.rm = TRUE))

norming_category_summary <- age_broad_category_predictions %>%
  group_by(is_norming) %>%
  summarize(overall_mean_prob = mean(mean_predicted_prob, na.rm = TRUE))

kable(summary(age_broad_category_predictions))
```

## Predicted Probabilities by Age

@fig-agepercent illustrates the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) overlaid to highlight the variability in predictions. Based on the resample of the test data, the scatterplot points represent individual predicted probabilities, while the percentile lines depict the progression of predictions across age groups.

The median predicted probability (50th percentile, solid blue line) steadily increases with age, reflecting the model’s growing confidence in high vocabulary acquisition as children get older. The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable rise after 20 months, suggesting greater variability in predictions among younger children. In contrast, the 75th and 90th percentiles (dashed and dotted purple/red lines) start higher and climb more sharply, indicating that some children exhibit advanced vocabulary skills even at younger ages.

Overall, the increasing spread between the percentile lines with age highlights a broader range of vocabulary acquisition patterns as children develop. This visualization underscores the model’s ability to account for individual differences in learning trajectories while confirming the strong association between age and predicted probability of high vocabulary.

```{r}
#| echo: false
#| eval: true
#| label: fig-agepercent
#| warning: false

# Resample the dataset to include 800 observations
set.seed(666)  # For reproducibility
sampled_data_test <- analysis_data_test %>%
  sample_n(5000)

# Define age ranges
sampled_data_test <- sampled_data_test %>%
  mutate(age_range = cut(age, breaks = seq(min(age), max(age), by = 3), include.lowest = TRUE, right = FALSE))

# Calculate percentiles for each age range
percentile_data <- sampled_data_test %>%
  group_by(age_range) %>%
  summarize(
    p10 = quantile(predicted_prob, 0.10, na.rm = TRUE),
    p25 = quantile(predicted_prob, 0.25, na.rm = TRUE),
    p50 = quantile(predicted_prob, 0.50, na.rm = TRUE),  # Median
    p75 = quantile(predicted_prob, 0.75, na.rm = TRUE),
    p90 = quantile(predicted_prob, 0.90, na.rm = TRUE),
    age_mid = mean(as.numeric(as.character(age)), na.rm = TRUE), # Midpoint of age range
    .groups = "drop")

# Create the scatterplot with percentile lines
ggplot(sampled_data_test, aes(x = age, y = predicted_prob)) +
  # Scatterplot of individual predictions
  geom_point(alpha = 0.5, size = 2, color = "#FF9933") +
  # Percentile lines
  geom_line(data = percentile_data, aes(x = age_mid, y = p10), 
            color = "#FF6666", linetype = "dashed", size = 1.2, show.legend = TRUE) +  # 10th percentile
  geom_line(data = percentile_data, aes(x = age_mid, y = p25), 
            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 25th percentile
  geom_line(data = percentile_data, aes(x = age_mid, y = p50), 
            color = "#3399FF", linetype = "solid", size = 1.5) +  # 50th percentile (Median)
  geom_line(data = percentile_data, aes(x = age_mid, y = p75), 
            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 75th percentile
  geom_line(data = percentile_data, aes(x = age_mid, y = p90), 
            color = "#FF6666", linetype = "dashed", size = 1.2) +  # 90th percentile
  # Add labels and theme
  labs(
    title = "Predicted Probabilities by Age with Percentiles", 
    x = "Age (Months)",
    y = "Predicted Probability",
    caption = "Percentile lines show 10th, 25th, 50th, 75th, and 90th percentiles."
  ) +
  scale_x_continuous(
    breaks = seq(min(sampled_data_test$age), max(sampled_data_test$age), by = 3),
    labels = function(x) paste(x, "mo")  # Add "mo" for months
  ) +
  scale_y_continuous(
    limits = c(0, 1), breaks = seq(0, 1, 0.1),
    labels = scales::percent_format()  # Format as percentages
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.caption = element_text(size = 10, hjust = 0.5, face = "italic"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank()) +
  annotate("segment", x = max(sampled_data_test$age) - 4, xend = max(sampled_data_test$age) - 2, y = 0.9, yend = 0.9, 
           color = "#FF6666", linetype = "dashed", size = 1.2)
```

## Distribution of Predicted Probabilities by Word Category

The distribution of predicted probabilities for achieving high vocabulary varies across word categories, as shown in @fig-prediction. Categories such as Function Words, Living Things, and Objects display a concentration of higher predicted probabilities, indicating consistent acquisition patterns and the foundational role of these words in early communication. In contrast, categories like Sensory Words and Verbs show a higher density of lower probabilities, reflecting their later emergence in language development and the contextual complexity they require.

Categories such as Adjectives and Places exhibit broader distributions, spanning a wider range of predicted probabilities. This variability highlights differences in children’s exposure to and use of these words, suggesting that external factors like linguistic environments may play a significant role. Overall, these results emphasize how word categories shape the model’s predictions and highlight the importance of addressing variability in complex word types for targeted interventions.

```{r, fig.width=10, fig.height=10}
#| echo: false
#| eval: true
#| label: fig-prediction
#| fig-cap: This figure displays the density distributions of predicted probabilities for achieving high vocabulary across different word categories. Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns. While categories like Verbs and Sensory Words exhibit broader distributions and lower probabilities, highlighting their greater complexity and variability in early language development.
#| warning: false

ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
  geom_density(alpha = 0.7, color = "black", size = 0.4) +  # Smooth density lines
  facet_wrap(~ broad_category, scales = "fixed", ncol = 4) +  # Use 2 columns for less crowding
  labs(
    title = "Distribution of Predicted Probabilities by Word Category",
    x = "Predicted Probability",
    y = "Density"
  ) +
  scale_fill_brewer(palette = "Pastel1") +  # Subtle and cohesive color palette
  theme_minimal(base_size = 15) +  # Increase base font size for better readability
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),  # Larger title
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),  # Increase axis text size
    strip.text = element_text(size = 14, face = "bold", color = "darkblue"),  # Larger, bold facet labels
    panel.spacing = unit(1.5, "lines"),  # Moderate spacing between facets
    legend.position = "none",  # Remove legend to declutter
    panel.grid.major = element_line(color = "gray85"),  # Light gray gridlines
    panel.grid.minor = element_blank()  # No minor gridlines for a clean look
  )

```

# Discussion {#sec-discussion}

## Age and Developmental Trajectories

The findings emphasize age as the most influential predictor of vocabulary acquisition. The sharp increase in predicted probabilities as children grow older aligns with expectations, as the developmental window between 18 and 30 months is crucial for rapid language learning. The steady upward trend in median vocabulary probabilities suggests that this period is marked by consistent growth in both production and comprehension. However, the growing variability between the upper and lower percentiles reveals the diverse learning trajectories among children. While some children quickly advance, others may require more time, reflecting the role of environmental factors and individual differences in cognitive development.

Interestingly, the plateauing trend observed in certain facets of comprehension suggests that as children master foundational vocabulary, their acquisition of more complex or context-specific words becomes less predictable. This shift highlights the importance of contextual factors such as exposure to rich language environments and interactions with caregivers.

## Lexical Categories: Foundation and Complexity

The analysis of word categories revealed striking differences in acquisition patterns. Categories such as “Function Words,” “Living Things,” and “Objects” displayed higher predicted probabilities, likely because these words are more frequently encountered in everyday interactions and form the backbone of early communication. In contrast, categories like “Verbs” and “Sensory Words” demonstrated lower probabilities and wider variability, suggesting they are more challenging for young learners. These words often require greater contextual understanding or abstract thinking, which may develop later in the language acquisition process.

The variability in word categories sheds light on how children’s experiences shape their vocabulary. For instance, categories with higher predicted probabilities may reflect a shared baseline of language exposure across children, whereas the broader spread in categories like “Adjectives” or “Sensory Words” points to the influence of specific cultural or environmental factors.

The acceleration in early vocabulary is even clearer when looking at production reports from older children using Words & Sentences. Figure 5.3 shows this pattern. In every language, the median child is reported to produce 50 words between 16–20 months (dotted line), though – as we will see below – this analysis masks tremendous between-child variability during this period. In addition, languages vary considerably in the absolute number of words reported. (As it is a major outlier, we have discussed the Beijing Mandarin WS data in Chapter 3, section on difficult data). Nevertheless, there are still substantial consistencies in the shape and general numerical range across languages.

During the period of 24–30 months, we see curves leveling out. Presumably, this leveling does not reflect a slowing in the rate of acquisition, which most researchers assume continues unabated for many years (e.g., Bloom, Tinker, and Scholnick 2001). Instead, it reflects the limitations of the CDI instrument, in that there are many possible “more advanced” words that children are likely learning, of which only a small subset are represented on any form.

The results highlight distinct trajectories in vocabulary development among children. Production scores showed a steady increase with age, with notable accelerations between 18 and 30 months. This period is critical for vocabulary acquisition, as children rapidly expand their linguistic capabilities. The variability observed in percentile distributions emphasizes individual differences in learning rates, influenced by factors such as exposure to language, socio-economic background, and cognitive abilities.

The analysis of comprehension data revealed a similar trend, with a marked rise in vocabulary comprehension after 18 months. Median comprehension scores provided a robust central tendency, effectively avoiding skewness caused by outliers. The plateau observed around 30 months suggests a stabilization in vocabulary growth, likely reflecting the limitations of the CDI dataset rather than a developmental plateau.

## Broader Implications for Language Development

The findings of this study offer practical applications for early education and language interventions by highlighting specific word categories that are more challenging for children to acquire. For instance, verbs and sensory words, which often require greater contextual understanding, could be emphasized in storytelling, play, and structured educational settings to help bridge the gap in these more complex categories. Tailored approaches in language development programs can address individual learning needs, with children demonstrating slower growth in foundational vocabulary benefiting from increased exposure to high-frequency words. Conversely, those excelling in comprehension may gain from activities introducing less common or context-specific terms, thereby expanding their linguistic repertoire.

Key predictors, including age, norming status, and lexical categories, significantly influenced vocabulary acquisition. Age was the strongest determinant of language growth, with older children showing higher vocabulary scores, consistent with developmental expectations. Additionally, children in the norming group exhibited superior vocabulary acquisition, potentially due to the structured linguistic environments associated with their assessment. Lexical categories provided further directions into vocabulary development, with objects and verbs showing higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication. In contrast, sensory words and adjectives displayed less consistent patterns, suggesting their acquisition may occur later as children’s linguistic abilities become more sophisticated.

## Limitations and Future Directions

@song also point out in their study on 264 typically developing Chinese children (145 boys and 119 girls) were included from a longitudinal study of language. While the findings offer valuable directions, several limitations should be acknowledged. First, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities. Future studies could complement CDI data with observational or experimental measures to enhance reliability. Second, the cross-sectional nature of the data limits the ability to track individual developmental trajectories. Longitudinal studies are needed to capture within-child variability and the dynamics of vocabulary growth over time.Familial factors and reading or language related cognitive skills were found to be associated with these developmental subgroups.

Additionally, the dataset we have now is limited on the word categories. Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition. These factors could provide a more comprehensive understanding of the mechanisms underlying linguistic development.

The strongest developmental inferences can be made by the examination of longitudinal data, in which children’s individual development is measured multiple times using the same instrument. Unfortunately, relatively little of our CDI data comes from this type of repeated administration. Figure 3.5 shows the number of administrations for particular languages that come from longitudinal datasets with a particular depth. There is a substantial amount of two-administration longitudinal data for several languages, but only a few have more than two observations for individual children.In general, this aspect of our data is a consequence of the fact that, for normative datasets, pure cross-sectional data collection is used to ensure statistical independence between datapoints. Thus, we must typically settle for using the large amount of available cross-sectional data to average out individual variability.

\newpage

\appendix

# Appendix {#sec-data2}

## Survey Design: The CDI Framework

The MacArthur-Bates Communicative Development Inventories (CDI) form the foundation of the Wordbank dataset used in this study. The CDI is a parent-report instrument that collects data on children’s vocabulary comprehension and production through structured surveys. This methodology allows for large-scale data collection across diverse populations, balancing cost-efficiency and practicality. However, as a parent-reported tool, the CDI is subject to several biases. For instance, parental interpretations of word comprehension or usage can vary, and factors like educational background, socio-economic status, and linguistic environment may influence the accuracy of responses.

The design of the CDI ensures comprehensive coverage of early vocabulary acquisition by categorizing words into broad lexical groups (e.g., Function Words, Verbs, Sensory Words). This categorization supports nuanced analysis but introduces challenges. For instance, frequently observed word types like Function Words may be more accurately reported than abstract categories like Sensory Words, leading to potential underestimation in less salient categories. These characteristics highlight the trade-off between the CDI’s extensive reach and the subjective variability in parental reporting.

## Sampling Framework

The @wordbankr dataset includes norming and non-norming samples, which together capture a broad spectrum of linguistic development. Norming samples are designed to represent a balanced, standardized population, serving as benchmarks for vocabulary acquisition. These data allow for cross-study comparisons and robust generalizations but may not fully capture real-world variability. In contrast, non-norming samples encompass a wider range of linguistic environments, including underrepresented groups. This dual approach helps investigate effects of contextual diversity on language acquisition but also introduces challenges in integrating findings from norming and non-norming subsets.

One critical consideration in observational datasets like Wordbank is the potential for sampling biases. For example, children from bilingual households or lower socio-economic backgrounds may be underrepresented in the norming sample, leading to an incomplete picture of vocabulary acquisition patterns. Future surveys could benefit from targeted recruitment strategies to ensure more comprehensive sampling of diverse populations.

## Observational Nature of the Data

The Wordbank dataset is observational, capturing vocabulary acquisition as it naturally occurs in diverse contexts. While this design enhances ecological validity, it limits causal inference. Factors such as parental education, linguistic exposure, and cultural norms may confound the observed relationships between predictors (e.g., age, norming status) and vocabulary outcomes. For instance, children in norming samples might have greater exposure to structured learning environments, inflating their vocabulary scores compared to non-norming counterparts.

To address these limitations, simulation methods were applied to assess how sampling variability influences model estimates. Random subsamples were drawn from the data, and the logistic regression model was refitted to evaluate the consistency of predictor effects. Key predictors like age and norming status remained stable, confirming the robustness of the model. However, coefficients for categories like Sensory Words and Places exhibited higher variability, reflecting their smaller sample representation and the challenges of capturing abstract vocabulary through parent-reported data.

## Connecting to the Literature

The CDI framework and its use in Wordbank align with best practices in developmental research, as outlined in studies from @mayor. The test-retest reliability of the CDI supports its validity, although the variability in parental reporting—especially for comprehension items—has been noted. Advanced statistical techniques like Item Response Theory (IRT) have been employed to evaluate the discriminative power of individual CDI items, revealing strengths in measuring commonly acquired words and limitations for abstract concepts. These findings inform the design of future surveys, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.

## Simulation and Recommendations

To further validate the findings, simulations could be expanded to incorporate potential sampling biases. For example, generating synthetic datasets that account for underrepresented groups (e.g., bilingual or low socio-economic households) could show how these populations might influence observed patterns. Additionally, calibration exercises—where parents undergo brief training on how to interpret survey items—could enhance the consistency of responses and reduce reporting variability.

This appendix underscores the importance of survey design, sampling methodology, and observational data analysis in understanding early vocabulary acquisition. By addressing inherent limitations and leveraging advanced methodologies, future research can refine the robustness and applicability of findings, ensuring a more inclusive representation of children’s linguistic development.

# Model details {#sec-model2}

## Model Summary

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: This table provides coefficients summary for the logistic model used in this paper.
#| warning: false

modelsummary(logistic_model)
```

## Diagnostics

### Confusion Matrix
```{r}
#| echo: false
#| warning: false
#| message: false

analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))

# Preprocess test data
analysis_data_test <- analysis_data_test %>%
  mutate(prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
    high_vocabulary = ifelse(prod_comp_mean > 300, 1, 0),  # Binary target variable
    age_scaled = scale(age)  # Standardize age
  ) %>%
  drop_na(prod_comp_mean)  # Drop rows with NA in prod_comp_mean
analysis_data_test$age_scaled <- as.numeric(analysis_data_test$age_scaled)

# Predict probabilities for test data
predicted_probabilities <- predict(logistic_model, newdata = analysis_data_test, type = "response")

# Predict classes based on probabilities
predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)

# Create a confusion matrix
confusion_matrix <- table(
  Actual = analysis_data_test$high_vocabulary,
  Predicted = predicted_classes
)

# Compute accuracy, sensitivity, specificity, and precision
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])  # True Positives / Actual Positives
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])  # True Negatives / Actual Negatives
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])  # True Positives / Predicted Positives

# Create a dataframe for metrics
metrics <- data.frame(
  Metric = c("Accuracy", "Sensitivity (Recall)", "Specificity", "Precision"),
  Value = c(accuracy, sensitivity, specificity, precision)
)

# Display the table using kable
kable(metrics, col.names = c("Metric", "Value"), format = "markdown", digits = 2)
```

### ROC Curve and AUC
```{r}
#| echo: false
#| warning: false
#| message: false

library(pROC)
roc_curve <- roc(analysis_data_test$high_vocabulary, predicted_probabilities)
auc_value <- auc(roc_curve)
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
```

### Precision-Recall Curve
```{r}
#| echo: false
# Precision-Recall Curve (optional, useful for imbalanced data)
#install.packages("PRROC")
library(PRROC)
pr_curve <- pr.curve(scores.class0 = predicted_probabilities[analysis_data_test$high_vocabulary == 1],
                     scores.class1 = predicted_probabilities[analysis_data_test$high_vocabulary == 0],
                     curve = TRUE)
plot(pr_curve, main = "Precision-Recall Curve")
```
### Binned Residual Plot
```{r}
#| echo: false

binnedplot(fitted(logistic_model), residuals(logistic_model, type = "response"),
           xlab = "Fitted Probabilities", ylab = "Average Residual",
           main = "Binned Residual Plot")
```

# Acknowledgements

This project was conducted under the help of OpenAI’s ChatGPT 4.0, which provided invaluable assistance in drafting and refining the paper. The analysis was conducted using a suite of packages from @citeR, which offered robust functionality for data manipulation, visualization, and storage. We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @arm, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes. Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files. Thanks to @proc and @prroc for supporting model diagnostics in this study.

A special acknowledgment goes to the @wordbankr team for providing the extensive dataset that forms the foundation of this research. Their contribution enabled a comprehensive exploration of vocabulary learning patterns in children. We are deeply grateful to the developers and maintainers of these open-source tools and datasets for their efforts in advancing research and accessibility in the data science community.

\newpage

# References